### 2013-06-28 Tex/chapters/monte_carlo.tex
Mehr erläuterungen zur autocorr zeit

\subsection{Equilibration- and Autocorrelation Time}
\label{ssec:eqtime}
    To generate an equilibrium state one starts with an arbitrary state
    and waits until it is equilibrated. The count of sweeps till
    equilibrium is called \emph{equilibration time} \(t_{eq}\).
    All measurement should start after this time.\\
    To determine when the system is in equilibrium, one can watch the
    development of some observables and take the point at which there
    are no big changes anymore as equilibrium. In fig.
    \ref{fig:equiandauto}\subref{sfig:equiandauto:equiE}
    the equilibrium is reached after approximately \(N=50\) sweeps for
    either an initial condition of all spins up and all spins random. It
    does not harm to double that value to be save.
    \begin{figure}[htbp]
        \centering
        \subfigure[Example of an equilibrating Ising system][]{
                \label{sfig:equiandauto:equiE}
                \includegraphics[width=0.47\textwidth]{plots/equiE}
        }
        \subfigure[Example of the autocorrelation of an Ising system][]{
                \label{sfig:equiandauto:autoM}
                \includegraphics[width=0.47\textwidth]{plots/autoM}
        }
        \caption[Examples for equilibration and autocorrelation]
        {
            \subref{sfig:equiandauto:equiE} Example of a Ising system
                \(L=64\) reaching equilibrium at \(T=2.36\) and
            \subref{sfig:equiandauto:autoM} the autocorrelation of an
                Ising system \(L=64\) at \(T=2.40\) (only Metropolis
                sweeps) on half logarithmic axis.
                The straight line is a exponential fit \(\exp(-t/\tau)\)
                with \(\tau = 342(1)\).
        }
        \label{fig:equiandauto}
    \end{figure}\\
    Because every state is generated from the state before, measurements
    of subsequent states are correlated. To determine when two states
    are independet, one calculates the normalized autocorrelation function
    \(\frac{\chi(t)}{\chi(0)}\), which should decay exponentially
    \(\chi(t) \propto \exp(t/\tau)\). This is visible in the half
    logarithmic plot \ref{fig:equiandauto}\subref{sfig:equiandauto:autoM}.
    To get the autocorrelation time one can either fit a exponential
    function \(\exp(-t/\tau)\) like in fig \ref{fig:equiandauto}\subref{sfig:equiandauto:autoM}
    or integrate \(\tau = \int \frac{\chi(t)}{\chi(0)} \de t\).
    \(\tau\) is an estimate after which time two samples are not
    correlated anymore. \cite[S. ??]{NewmanBarkema1999} \cite[S. 150f]{Katzgraber2011}.
    To make sure that the error is not underestimated one should wait
    \(2\tau\) sweeps between two measurements.

>>>>

\subsection{Equilibration- and Autocorrelation Time}
\label{ssec:eqtime}
    To generate an equilibrium state one starts with an arbitrary state
    and waits until it is equilibrated. The count of sweeps till
    equilibrium is called \emph{equilibration time} \(t_{eq}\).
    All measurement should start after this time.\\
    To determine when the system is in equilibrium, one can watch the
    development of some observables and take the point at which there
    are no big changes anymore as equilibrium. In fig.
    \ref{fig:equiandauto}\subref{sfig:equiandauto:equiE}
    the equilibrium is reached after approximately \(N=50\) sweeps for
    either an initial condition of all spins up and all spins random. It
    does not harm to double that value to be save.
    \begin{figure}[htbp]
        \centering
        \subfigure[Example of an equilibrating Ising system][]{
                \label{sfig:equiandauto:equiE}
                \includegraphics[width=0.47\textwidth]{plots/equiE}
        }
        \subfigure[Example of the autocorrelation of an Ising system][]{
                \label{sfig:equiandauto:autoM}
                \includegraphics[width=0.47\textwidth]{plots/autoM}
        }
        \caption[Examples for equilibration and autocorrelation]
        {
            \subref{sfig:equiandauto:equiE} Example of a Ising system
                \(L=64\) reaching equilibrium at \(T=2.36\) and
            \subref{sfig:equiandauto:autoM} the autocorrelation of an
                Ising system \(L=64\) at \(T=2.40\) (only Metropolis
                sweeps) on half logarithmic axis.
                The straight line is a exponential fit \(\exp(-t/\tau)\)
                with \(\tau = 342(1)\).
        }
        \label{fig:equiandauto}
    \end{figure}\\
    Because every state is generated from the state before, measurements
    of subsequent states are correlated. To determine when two states
    are independet, one calculates the normalized autocorrelation function
    \(\frac{\chi(t)}{\chi(0)}\), which should decay exponentially
    \(\chi(t) \propto \exp(t/\tau)\). This is visible in the half
    logarithmic plot \ref{fig:equiandauto}\subref{sfig:equiandauto:autoM}.
    To get the autocorrelation time one can either fit a exponential
    function \(\exp(-t/\tau)\) like in fig \ref{fig:equiandauto}\subref{sfig:equiandauto:autoM}
    or integrate \(\tau = \int \frac{\chi(t)}{\chi(0)} \de t\).
    \(\tau\) is an estimate after which time two samples are not
    correlated anymore. \cite[S. ??]{NewmanBarkema1999} \cite[S. 150f]{Katzgraber2011}.
    To make sure that the error is not underestimated one should wait
    \(2\tau\) sweeps between two measurements.
    The autocorrelation time is of course dependent on the temperature.
    At high temperatures the fluctuations are strong and subsequent
    states are more dissimilar and therefore less correlated than at low
    temperatures, where less spins are flipping. In this thesis always
    the longest autocorrelation time \(\tau\) of all simulated
    temperatures was taken.
    \[\tau = \max_T \tau_T\]

### 2013-06-28 Tex/chapters/results.tex
List of Tables für die Vollständigkeit
Mehr erläuterungen zur autocorr zeit

For evaluation the Monte Carlo simulation is run until the system
is equilibrated after \(t_{eq}\) sweeps. Then the simulation continues
and the magnetisation per site \(m=\frac{1}{N}\sum_i s_i\) and energy
per site \(E=\frac{1}{N}\hat H\) are calculated and saved for every
\(2\tau\) sweeps. Where \(\tau\) denotes the \emph{autocorrelation time}
(see also \ref{ssec:eqtime}). The used values for different system sizes
are listed in tab. \ref{tab:tauAndTeq} in numbers of sweeps. Note that
these values are generously rounded up to be on the safe side.
\begin{table}[htbp]
    \center
    \begin{tabular}{l r r r r r}
        \toprule
        \(L\)    & 16 &  32 &  64 & 128 & 256\\
        \midrule
        \(\tau\) &  3 &   5 &   7 &  12 &  16\\
        \(t_eq\) & 40 & 100 & 100 & 200 & 600\\
        \bottomrule
    \end{tabular}
    \caption{Autocorrelation Times $\tau$ and Equilibration Times $t_{eq}$}
    \label{tab:tauAndTeq}
\end{table}\\
For every observable \(O\) the expected value \(\avg{O}\) is determined
as the mean of \(10000\) measurements for \(L=16,32\) or \(5000\) for
\(L=64,128\). The expected values \(\avg{O}\) for \(100\) different
random proximity graphs with the same disorder paramter \(\sigma\)
are then averaged to \(\overline{\avg{O}}\). The errors are estimated
by bootstrapping \cite{Bootstrap}.

@@

\subsection{Critical Exponents}
    For \(\sigma \in \{0,0.1,0.2,0.3,0.5,1.0\}\) a finite size scaling analysis was
    performed to determine the critical exponents \(\beta, \gamma, \nu\)
    using \texttt{autoscale.py} \cite{autoscale2009}. The values for
    \(\sigma = 0\) are analytically known \cite{Pelissetto2002}. The
    values for all other \(\sigma\) should be the same as for \(\sigma = 0\)
    like mentioned before in section \ref{ssec:finitesize}.
    The values for \(\sigma\) are chosen to represent every interesting
    region from fig. \ref{fig:Tc}\subref{sfig:Tc_norm:RNG}\subref{sfig:Tc_norm:GG}:
    the plateau at small \(\sigma\), the steep decline and the plateau at
    big \(\sigma\).\\
    Like in tab. \ref{tab:critExp} to see, most values
    are matching the expectations. Most \(\beta\) seem to be a bit too
    big, but they are close enough to the expectations to be explained
    by the fact that small systems (\(L=32,64\)) were used for the
    analysis.
    Anyway, two critical exponents are sufficient to determine the universality
    class. Therefore this disordered Ising modell is in the same universality
    class as the standard Ising ferromagnet \cite[p. 145]{Katzgraber2011}.
    \begin{table}[htbp]
        \center
        \begin{tabular}{l l l l l}
            \toprule
             & \multicolumn{1}{c}{\(\sigma\)} & \multicolumn{1}{c}{\(\nu\)} & \multicolumn{1}{c}{\(\gamma\)} & \multicolumn{1}{c}{\(\beta\)}\\
            \midrule
            exact (\cite[p. 59]{Pelissetto2002}) & \multicolumn{1}{c}{\(0\)} & \multicolumn{1}{c}{\(1\)} & \multicolumn{1}{c}{\(-\frac{7}{4}\)} & \multicolumn{1}{c}{\(\frac{1}{8}\)}\\
            \midrule
            Gabriel      & 0.0 & 1.008(4) & -1.735(2) & 0.1262(4)\\
                         & 0.1 & 1.02(1)  & -1.744(5) & 0.133(6) \\
                         & 0.3 & 1.000(5) & -1.724(16)& 0.129(12)\\
                         & 0.5 & 1.009(8) & -1.750(12)& 0.125(13)\\
                         & 1.0 & 1.015(22)& -1.743(17)& 0.123(16)\\
            \midrule
            Relative N.  & 0.0 & 1.007(2) & -1.739(2) & 0.130(1) \\
                         & 0.1 & 0.99(1)  & -1.746(5) & 0.133(4) \\
                         & 0.2 & 1.022(17)& -1.756(14)& 0.123(10)\\
                         & 0.5 & 1.002(17)& -1.750(16)& 0.143(13)\\
                         & 1.0 & 1.011(20)& -1.758(16)& 0.138(13)\\
            \bottomrule
        \end{tabular}
        \caption{Critical exponents for different \(\sigma\)}
        \label{tab:critExp}
    \end{table}

>>>>

For evaluation the Monte Carlo simulation is run until the system
is equilibrated after \(t_{eq}\) sweeps. Then the simulation continues
and the magnetisation per site \(m=\frac{1}{N}\sum_i s_i\) and energy
per site \(E=\frac{1}{N}\hat H\) are calculated and saved for every
\(2\tau\) sweeps. Where \(\tau\) denotes the \emph{autocorrelation time}
(see also \ref{ssec:eqtime}). The used values for different system sizes
are listed in tab. \ref{tab:tauAndTeq} in numbers of sweeps. Note that
these values are generously rounded up to be on the safe side.
\begin{table}[htbp]
    \center
    \begin{tabular}{l r r r r r}
        \toprule
        \(L\)    & 16 &  32 &  64 & 128 & 256\\
        \midrule
        \(\tau\) &  3 &   5 &   7 &  12 &  16\\
        \(t_eq\) & 40 & 100 & 100 & 200 & 600\\
        \bottomrule
    \end{tabular}
    \caption{Autocorrelation Times $\tau$ and Equilibration Times $t_{eq}$}
    \label{tab:tauAndTeq}
\end{table}\\
Obiviously the autocorrelation time \(\tau\) increases with the system
size. In fact it obeys a power law as clearly visible in fig.
\ref{fig:autocorrPowerLaw}.
% sagen, dass das bekanntes verhalten ist
% als cross check behandeln
% [citation needed]
\begin{figure}[htbp]
    \centering
    \includegraphics[width=0.45\textwidth]{plots/autocorr}
    \caption
    {
        Dependece of the autocorrelation time $\tau$ on the system size $L$
    }
    \label{fig:autocorrPowerLaw}
\end{figure}\\
For every observable \(O\) the expected value \(\avg{O}\) is determined
as the mean of \(N_{\mathrm{measure}}=10000\) measurements for \(L=16,32\)
or \(N_{\mathrm{measure}}=5000\) for \(L=64,128\). The number of
calculated sweeps totals to
\[N_{\mathrm{sweeps}}=t_{eq}+2\tau N_{\mathrm{measure}}\]
The expected values \(\avg{O}\) for \(100\) different random proximity
graphs with the same disorder paramter \(\sigma\) are then averaged to
\(\overline{\avg{O}}\).
The errors are estimated by bootstrapping \cite{Bootstrap}.

@@

\subsection{Critical Exponents}
    For \(\sigma \in \{0,0.1,0.2,0.3,0.5,1.0\}\) a finite size scaling analysis was
    performed to determine the critical exponents \(\beta, \gamma, \nu\)
    using \texttt{autoscale.py} \cite{autoscale2009}. The values for
    \(\sigma = 0\) are analytically known \cite{Pelissetto2002}. The
    values for all other \(\sigma\) should be the same as for \(\sigma = 0\)
    like mentioned before in section \ref{ssec:finitesize}.
    The values for \(\sigma\) are chosen to represent every interesting
    region from fig. \ref{fig:Tc}\subref{sfig:Tc_norm:RNG}\subref{sfig:Tc_norm:GG}:
    the plateau at small \(\sigma\), the steep decline and the plateau at
    big \(\sigma\).\\
    Like in tab. \ref{tab:critExp} to see, most values
    are matching the expectations. Most \(\beta\) seem to be a bit too
    big, but they are close enough to the expectations to be explained
    by the fact that small systems (\(L=32,64\)) were used for the
    analysis.
    Anyway, two critical exponents are sufficient to determine the universality
    class. Therefore this disordered Ising modell is in the same universality
    class as the standard Ising ferromagnet \cite[p. 145]{Katzgraber2011}.
    \begin{table}[htbp]
        \center
        \begin{tabular}{l l l l l}
            \toprule
             & \multicolumn{1}{c}{\(\sigma\)} & \multicolumn{1}{c}{\(\nu\)} & \multicolumn{1}{c}{\(\gamma\)} & \multicolumn{1}{c}{\(\beta\)}\\
            \midrule
            exact (\cite[p. 59]{Pelissetto2002}) & \multicolumn{1}{c}{\(0\)} & \multicolumn{1}{c}{\(1\)} & \multicolumn{1}{c}{\(-\frac{7}{4}\)} & \multicolumn{1}{c}{\(\frac{1}{8}\)}\\
            \midrule
            Gabriel      & 0.0 & 1.008(4) & -1.735(2) & 0.1262(4)\\
                         & 0.1 & 1.02(1)  & -1.744(5) & 0.133(6) \\
                         & 0.3 & 1.000(5) & -1.724(16)& 0.129(12)\\
                         & 0.5 & 1.009(8) & -1.750(12)& 0.125(13)\\
                         & 1.0 & 1.015(22)& -1.743(17)& 0.123(16)\\
            \midrule
            Relative N.  & 0.0 & 1.007(2) & -1.739(2) & 0.130(1) \\
                         & 0.1 & 0.99(1)  & -1.746(5) & 0.133(4) \\
                         & 0.2 & 1.022(17)& -1.756(14)& 0.123(10)\\
                         & 0.5 & 1.002(17)& -1.750(16)& 0.143(13)\\
                         & 1.0 & 1.011(20)& -1.758(16)& 0.138(13)\\
            \bottomrule
        \end{tabular}
        \caption{Critical exponents for different $\sigma$}
        \label{tab:critExp}
    \end{table}
